A scenario for a Planckian standard kilogram

### 2017-10-30 manuscript_gravitationalconstant.tex
corrections

\begin{abstract}
The kilogram could, in principle, be redefined as an integer multiple
of the Planck mass (or 4594xxxx$\times\sqrt{\hbar c/G}$) with no
significant change from its current value.  The practical hurdle is
measuring $G$ three or four orders of magnitude than presently known,
so as to find the unknown digits xxxx.  We study a possible scenario
for doing so with an artificial two-body orbit.  Requirements for
suppressing perturbing effects (notably tidal forces and radiation
pressure), and obtaining the desired signal-to-noise are estimated.
\end{abstract}

@@

The reform of SI units to use quantum standards is an ongoing `news
story' in physics.  At the heart of the reform are two quantum
phenomena that appear on macroscopic scales: the Josephson effect
makes $h/e$ measurable as a voltage-to-frequency ratio, while the
quantum Hall effect makes $h/e^2$ a standard resistance (the
von~Klitzing constant).  These enable electrical units to be defined
in terms of $h$ and $e$.  A further ingenious instrument, known as a
watt balance, precisely compares electrical and mechanical power, thus
enabled mass to be standardized in terms of $h$.  For an simple
introduction to the new SI, we refer the reader to the article on the
``LEGO watt balance''[2015AmJPh..83..913C]\cite{Chao15}.

@@

Historically, there have been several proposed unit systems that
included $G$ in some way,\cite{Tomilin1999} but the idea of including
both gravity and quantum mechanics in a unit system comes from
Planck\cite{Planck99}.  In the process of working towards the
radiation law, Planck realized that $c,G,h$ and the Boltzmann constant
$k_B$ (in his notation $G\rightarrow f$, $h\rightarrow b$,
$h/k_B\rightarrow a$) naturally define units for mass, length, time,
and temperature.  (Planck remarked that such units would be meaningful
even in non-human and extraterrestrial cultures.)

A hypothetical Planckian SI unit system would have
\begin{equation} \label{eq:defs}
\begin{aligned}
\mathrm{kg} &= 4.594\ldots \times 10^7
               \left(\frac{\hbar c}{G}\right)^{\!1/2} \\
\mathrm{m}  &= 6.187\ldots \times 10^{34}
               \left(\frac{\hbar G}{c^3}\right)^{\!1/2} \\
\mathrm{s}  &= 1.854\ldots \times 10^{43}
               \left(\frac{\hbar G}{c^5}\right)^{\!1/2} \\
\mathrm{K}  &= 7.058\ldots \times 10^{-33}
               \left(\frac{\hbar c^5}{Gk_B^2}\right)^{\!1/2}
\end{aligned}
\end{equation}
with the numerical coefficients chosen so as to match the present
standards.  The square-rooted expressions are conventionally known as
the Planck mass, Planck length, and so on.  (They are actually
$\sqrt{2\pi}$ times Planck's definitions.)  Extension to electrical
units is straightforward, but we will not discuss it here.  The Planck
length, time, and temperature in themselves are extreme scales, not
probed by any known physical process, and presumably the regime of
quantum gravity.  But extreme scales do not harm if they are only part
of some expression, as they are here.  The Planck mass is small but
microscopically measurable: $\sim20\mu\mathrm{g}$.  This is of the
same order as the uncertainty in the prototype kilogram itself.  Hence
the kilogram could be redefined as an integer multiple of the Planck
mass, without significantly perturbing its value.

The difficulty with Eq.~\eqref{eq:defs} is that all the expressions
involve $G$.  Given the current uncertainties in $G$,
Eq.~\eqref{eq:defs} would allow units to be standardized only to four
significant digits, which is obviously not enough.  Improvement by
several orders of magnitude would be needed, to make the Planckian
kilogram practicable.  In this paper, we study a scenario which may
make such an improvement possible.  A Planckian second not a prospect
with anything like present technology, because demands on the time
standard are more stringent still.\\

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{img/plotGmeas}
	\caption{The data points represent six of the most recent measurement of the gravitational constant. The green line represents the recommended CODATA value and its uncertainty.}
	\label{fig:Gmeasurements}
\end{figure}

@@

In Fig.~\ref{fig:Gmeasurements} we have plotted some of the most recent measurements of the gravitational constant. Three of them have been carried out using a torsion balance (BIPM-14\cite{Quinn14}, UCI-14\cite{Newman14}, HUST-09\cite{Luo09}), UZur-06\cite{Schlamminger06} used a beam balance and JILA-10\cite{Parks10} measured the separation of two simple pendulums in a gravitational field. They all tried to measure the very weak gravitational force with a mechanical apparatus. The sixth measurement (LENS-14\cite{Rosi14}) was performed under another approach, using cold atom interferometry. Its uncertainty is a lot larger than the others, but it could become more competitive in the future. However, a brief glance on the plotted measurements shows the disagreement of current $G$-values and suggests a demand for new, potentially different techniques.\\

\subsection{Description of new measurement}
Newton's law of gravitation does not only explain why an apple falls down from a tree, it also explains orbits. If one has two bodies of masses $m_1$ and $m_2$, on which the only force acting is their respective gravitational attraction, they will orbit around their common barycenter in an ellipse described by Kepler's laws. The third law, given in Eq.~(\ref{eq:kepler3}), states that the square of the orbital period $P$ is directly proportional to the cube of the semimajor axis $a$. In this proportionality factor, the gravitational constant $G$ appears.
\begin{equation}\label{eq:kepler3}
P^2 = \frac{4 \, \pi^2}{G(m_1+m_2)}a^3
\end{equation}
If one can measure the position of the two masses very accurately and thus knows the orbit, it is possible to deduce the gravitational constant. 
So what we need is a freely falling two-body system. The only way to achieve this over some time, is to go to space. Thus, one has to design a spacecraft involving a cavity that contains the two masses and a position measurement apparatus. The cavity should shield the two-body system from environmental influences. The spacecraft is then put in space, preferably to a Lagrange point of the earth-sun system, in order to prevent large tidal forces.\\

@@

To get a feeling what the dimensions of such an experiment could be, choose for example lead balls with a diameter of about 10 cm, their mass then is about 6 kg. If one then chooses a semimajor axis of 15 cm, the orbital period would be roughly 3.5 hours and the gravitational force acting between the two bodies is of order $10^{-8}$ N. In Fig.~\ref{fig:sketch} we show a sketch on how this would look like.
For getting an improvement in the precision of the gravitational constant, the positions of the two balls have to be known very accurately. Best suitable would probably be a laser interferometry system, which can yield precisions down to some nanometers.\cite{Loughridge13}
The main difference to other techniques is that position measurements can be very accurate compared to directly measuring the gravitational force.
This new kind of measurement of the gravitational constant appears simple, but it brings various difficulties with it. It is practically impossible to have a clean system, where only the respective gravitational force is acting between the two bodies, so one has to find out what effects could possibly disturb the system and then either make them as small as possible or correct for them. \\

@@

\textbf{Radiation pressure:} The envelope of the cavity needs to protect the two body-system from the constant bombardment of radiation and charged particles from the sun. Since the activity of the sun varies much, the spacecraft must be able to correct its position around the freely falling masses inside the cavity.\\
Furthermore, there is thermal radiation inside the cavity, that could disturb the orbit.
Using the Stefan-Boltzmann law, we can relate radiation pressure to the temperature inside the cavity. If we then specify to balls of radius $r$, we can express the force on the balls as 
\begin{equation}
F = \frac{8 \pi^6 r^2 {k_\mathrm{B}}^4 T^4}{45 c^3 h^3}.
\end{equation}
%
%To estimate the force on a ball exerted by radiation, we use the Stefan-Boltzmann law. The radiation pressure $p_r$ can be written as a third of the energy density of the electromagnetic radiation and with fundamental thermodynamics one can relate the radiation pressure to the temperature $T$
%
%\begin{equation}
%\label{Stefan-Boltzmann}
%p_r = \frac{1}{3} u = \frac{1}{3} a T^4
%\end{equation}
%where the constant $a$ follows from Planck's radiation law and is
%\begin{equation}
%\label{eq:a}
%a = \frac{8 \pi^5 {k_\mathrm{B}}^4}{15 c^3 h^3}.
%\end{equation}
%We have now a direct relation between radiation pressure and the temperature in the cavity. Then we can estimate the force exerted on a ball of radius $r$ by just multiplying the pressure with the area
%\begin{equation}
%F = \frac{1}{3} \pi r^2 a T^4.
%\end{equation}
%
As we can see, this force is highly temperature dependent. In order to minimize this influence, one should deploy a temperature as small as possible in the cavity. In fact, if a cooling system based on liquid helium is installed, temperature can be brought down to about 4 K. Such a low temperature results in a force on the balls in the order of $10^{-16}$ N. Compared to the gravitational force, this is orders of magnitude less and thus we don't have to worry about effects of radiation pressure.\\

\textbf{Influence of the position measurement:} 
The photons of the laser that measures the position of the two balls exert an additional force. If a laser with a power of 1 mW is used and not more than five measurements every second are performed, the force on the balls is in the same order as the influence of thermal radiation and thus negligible too.\\
%
%In general, one has to pay attention as soon as one undertakes a measurement on a physical system, because every measurement can disturb the system. In this case we have a laser that measures the position of two masses which are in an orbit but the photons from the laser exerts an additional force on the masses. To give a rough estimation of this influence, we can compare the radiation pressure of the laser with the one of thermal radiation in the cavity, which we estimated before. In terms of the power $P$ of the laser and the area of the balls $A$, the radiation pressure is
%\begin{equation}
%\label{eq:radpress}
%p_r = \frac{P}{c\,A}.
%\end{equation}
%On behalf of minimizing the impact of the position measurement, a laser with small power should be used. If we assume a laser with a power of 1 mW, the radiation pressure is in the order of $10^{-10} \, \mathrm{N} \, \mathrm{m}^{-2}$. This is a factor of $10^4$ more than radiation pressure of thermal radiation in the cavity, but the laser doesn't have to shoot photons on the balls constantly. Only for the actual measurement a short pulse has to be sent. If this pulse has a duration of less than roughly $0.1$ ms, its disturbing effect on the orbit is in the same range as the effect of thermal radiation and is thus negligible too.\\
%
\textbf{Coulomb force:}
Since the electromagnetic force is orders of magnitude stronger than the gravitational force, the balls must be absolutely neutral. A possible source of charges are cosmic rays or the solar wind, so they have to be protected as good as possible. If we want that the Coulomb force between the balls is in the same order as the effect of thermal radiation and thus negligible, the maximal allowed charge on the balls is about $6 \cdot 10^{-14}\,\mathrm{C}$, which corresponds to roughly $4 \cdot 10^5$ elementary charges.\\

@@

\subsection{Orbital equation of motion}
From Newtonian dynamics the equation of motion for a clean two-body system is given by 
\begin{equation}
\label{eq:eomeasy}
\ddot{\vec{r}} = - \frac{G M\, \vec{r}}{r^3}.
\end{equation}
Here $M$ denotes the sum of the masses $m_1$ and $m_2$, $G$ is the gravitational constant and $r$ is the relative distance between the two bodies. For simplicity we will abbreviate the product of $GM$ with $k$.
Since out two-body system is in a rotating frame we have to add the term developed before for the Coriolis acceleration as well as a correction term proportional to the distance accounting for centrifugal and tidal forces. This last term can be written with a
symmetric $3 \times 3$ matrix $C$ as follows  
\begin{equation}
\label{eq:tidal}
\ddot{\vec{r}}_{\mathrm{C}} = C \cdot \vec{r}.
\end{equation}
or in terms of the individual components 
\begin{equation}
\label{eq:tidacom}
\left(\ddot{r}_\mathrm{C}\right)_i = \sum_{k=1}^{3} C_{ik} \: r_k.
\end{equation}
It can be thought of as a first order expansion and therefore also include other possible minor perturbations. The reference frame is chosen such that the barycenter of the two-body system is force-free.
So the full equation of motion for the two balls looks as follows:
\begin{equation}
\label{eq:eom}
\ddot{\vec{r}} = - \frac{k\,\vec{r}}{r^3} - 2 \; \vec{\omega} \times \dot{\vec{r}} + C \cdot \vec{r}
\end{equation}
These are actually three coupled equations of motion, one for each coordinate axis. Since all are second order differential equations, this leaves us with six initial conditions. In other words to determine the orbit, we need to give six starting values.\\
We choose our coordinate system such that the orbital plane of the two-body system is in the $x-y$ plane. $\vec{\omega}$ is a constant vector that points in the direction of the rotation axis of the earth-sun reference frame. We will fit for $\vec{\omega}$, so we don't have to know it exactly before, but it can be estimated well and thus provides good starting points for the fit. The same is valid for the six components of the matrix $C$.\\

@@

Our problem has 16 unknown parameters, but the only one we are really interested in is $k$, which contains the gravitational constant. For the fit, we suggest to use a Markov Chain Monte-Carlo (MCMC) method because it does good work even if the parameter space is large. We will not describe in detail how the MCMC is working, rather refer to an online tutorial from ESO PyCoffee\cite{esopycoffee} or an introduction paper from van Ravenzwaaij et al.\cite{vanRavenzwaaij16} \\

A Markov chain Monte-Carlo method is a Bayesian approach to data analysis. It uses the information provided by observed data about a set of parameters, to update a prior state of beliefs about the set of parameters to become a posterior state of beliefs about a set of parameters. First we have to specify the prior probability distribution, so to say what values for the 16 parameters in the equation of motion actually make sense. Then, for all 16 parameters a uniform distribution within these ranges and zero probability everywhere else is assumed as the prior probability distribution. \\

We briefly explain in what range the values of the parameters are estimated and display it in table~\ref{tab:parameters}.
For the parameter $k$, which is the product of the gravitational constant $G$ and the total mass $M$ we look at the current measured values for the gravitational constant in Fig.~\ref{fig:Gmeasurements}, it is reasonable to constrain its numerical value in SI units to values between $6.671 \cdot 10^{-11}$ and $6.676 \cdot 10^{-11}$. The masses of the two balls will not contribute to a larger uncertainty in $k$, we choose them to be 6 kg.\\
The vector $\vec{\omega}$ accounts for the rotation of the reference frame. If we say that the orbit of the balls lies in the same plane as the orbital plane of the earth, $\vec{\omega}$ will point in $z$ - direction with $x$ and $y$ components small. Thus $\omega_z$ is associated with the period $P_\odot$ of one year through $2\pi/P_\odot$. Because the rotation axis are conceivably not exactly aligned, we add some uncertainty. As mentioned before, the $x$ and $y$ components are small, so to say roughly in the range of the uncertainty in $w_z$.\\
The main contribution to the tidal term goes with $\omega^2$ and lies in the orbital plane. We are free to choose the $x$ and $y$ axis, so we choose it such that it is in $x$-direction. The other five components then should be smaller, but we give some extra range to account for possible other perturbations.\\
% 
%To estimate the six components of the matrix $C$, we first state again that the barycentre of the two balls is in free fall. So at this point all forces cancel out. If we imagine the ball on its orbit, the tidal force is stronger when its nearer to the earth-sun barycentre and the centrifugal force is stronger when it is in the opposite position. So over an orbit, the effect on the balls stays constant. Also other perturbations are assumed to be constant, at least during some orbits.
%The main component of this term should be perpendicular to the rotation axis, for which we have chosen the $z$ - axis. Since we are free to set the $x$ and $y$ axis in the orbital plane, we can choose it such that the main contribution lies in the $x$ - axis. Its magnitude goes as $\omega^2$, so the $C_{xx}$ component should be around $4 \cdot 10^{-14}$. Because we assumed that the main effect is in the $x$ - direction, all the other components are smaller, we constrain them to somewhere between $\pm 10^{-14}$. Since the $C$ matrix also absorbs other perturbations, we give some more uncertainty here. \\
%
The initial values for the positions can be guessed very accurately, since they are directly measured precisely. The initial velocity can be calculated easily and matched to the initial positions, therefore also yielding narrow parameter ranges.\\

In the next step, we set the parameters to a fixed ``true" value within the range estimated. We have to keep in mind that $k$, $x_0$ and $v_{y_0}$ are linked together and there must not be contradictions in the true values. The chosen values are displayed in the right column in table~\ref{tab:parameters}.\\

@@

To simulate data we would obtain from the position measurements, we integrate the equations of motion. The time for one orbit without perturbations can be estimated using Kepler's third law (\ref{eq:kepler3}). For the value of $k$ assumed above and a semi-major axis of $0.15 \, \mathrm{m}$, this results in an orbital period of roughly $12900 \, \mathrm{seconds}$ or $3.5 \, \mathrm{hours}$. Here we integrate two orbits in $100000$ steps, that would correspond to about $4$ measurements per second, which should be feasible. To the generated data, we add normal distributed noise with a standard deviation of $10 \, \mathrm{nm}$. This uncertainty stays for a systematic error in the laser interferometry measurement and also random statistic noise. Its value is just an estimation, it would depend highly on the measurement system that has been used.\\

@@

A useful way to look at the samples of the posterior returned by the MCMC is doing a corner plot using a Python module named \textit{corner}\cite{corner}. The corner plot shows all the one and two dimensional projections of the posterior probability distributions and demonstrates quickly the covariances between parameters. Along the diagonal, the marginalized distribution for each parameter independently is shown.
In Fig.~\ref{fig:plotCorner4} an excerpt of the full corner plot is shown. \\
%
%Here we only depict the parameters $x_0$, $v_{y_0}$, $\omega_z$, $C_{xx}$ and $k$. These are the more important ones, because for example the main contribution to $\vec{\omega}$ comes from its $z$-component.\\
%

To arrive at a best value with uncertainty for the gravitational constant, we take the mean and standard deviation of the marginalized distribution of the parameter $k$ and divide by the total mass of the balls. In this example the relative uncertainty of $k$ is roughly $10^{-7}$, so if the masses are measured to about the same precision, we can write for the gravitational constant $G$:
\begin{equation}\label{Gfit}
G = \left(6.674005 \pm 0.000001 \right) \cdot 10^{-11}  \frac{\mathrm{m}^3}{\mathrm{kg \, s}^2}
\end{equation}
This corresponds to a relative uncertainty of $1.5 \cdot 10^{-7}$ or 15 parts in a hundred million.\\

%
%we take the median and the percentiles corresponding to one sigma from the distribution of the parameter $k$. For this example we get:
%\begin{equation}\label{eq:kfit}
%k = \left( 8.0088006 \, \substack{+0.0000005 \\ -0.0000008}\, \right) \cdot 10^{-10}
%\end{equation}
%Now we divide $k$ by the total mass of the two balls. Note, that we assume, that the uncertainty on the mass of the balls does not contribute to a larger uncertainty on $G$, because their mass should be determinable to a relative uncertainty of about $10^{-8}$, whereas the one of $k$ is roughly $10^{-7}$. We change the way to write the uncertainty from the 0.16 and 0.84 percentiles to the usual standard deviation, while adding some margin. With units put back in, we arrive at the following value for the gravitational constant $G$:
%\begin{equation}\label{Gfit}
%G = \left(6.674005 \pm 0.000001 \right) \cdot 10^{-11}  \frac{\mathrm{m}^3}{\mathrm{kg \, s}^2}
%\end{equation}
%This corresponds to a relative uncertainty of $1.5 \cdot 10^{-7}$ or 15 parts in a hundred million.\\
%

\section{Conclusion}
We have proposed a new way to measure the gravitational constant $G$ and estimated on what order of magnitude different disturbing effects occur and if one has to correct for them. We then simulated measurement data and showed how to extract the gravitational constant. Based on our estimations, it should be possible to measure the gravitational constant to about $10^{-7}$ relative uncertainty. This is a about a factor of 300 better than today's recommended CODATA value.\\
If we are a bit optimistic, a further increase in precision is not out of reach, so that about the same relative uncertainty as for Planck's constant $h$ ($1.2 \cdot 10^{-8}$)\cite{CODATA16} could be achieved. This then would allow to redefine the kilogram as an integer multiple of the Planck mass.\\

@@

\begin{acknowledgments}
Need acknowledgments?
\end{acknowledgments}

@@

>>>>

\begin{abstract}
The kilogram could, in principle, be redefined as an integer multiple
of the Planck mass (or 4594xxxx$\times\sqrt{\hbar c/G}$) with no
significant change from its current value.  The practical hurdle is
measuring $G$ three or four orders of magnitude more precisely than presently known,
so as to find the unknown digits xxxx.  We study a possible scenario
for doing so with an artificial two-body orbit.  Requirements for
suppressing perturbing effects (notably tidal forces and radiation
pressure), and obtaining the desired signal-to-noise are estimated.
\end{abstract}

@@

The reform of SI units to use quantum standards is an ongoing `news
story' in physics.  At the heart of the reform are two quantum
phenomena that appear on macroscopic scales: the Josephson effect
makes $h/e$ measurable as a voltage-to-frequency ratio, while the
quantum Hall effect makes $h/e^2$ a standard resistance (the
von~Klitzing constant).  These enable electrical units to be defined
in terms of $h$ and $e$.  A further ingenious instrument, known as a
watt balance, precisely compares electrical and mechanical power, thus
enabled mass to be standardized in terms of $h$.  For a simple
introduction to the new SI, we refer the reader to the article on the
``LEGO watt balance''[2015AmJPh..83..913C].\cite{Chao15}

@@

Historically, there have been several proposed unit systems that
included $G$ in some way,\cite{Tomilin1999} but the idea of including
both gravity and quantum mechanics in a unit system comes from
Planck.\cite{Planck99}  In the process of working towards the
radiation law, Planck realized that $c,G,h$ and the Boltzmann constant
$k_B$ (in his notation $G\rightarrow f$, $h\rightarrow b$,
$h/k_B\rightarrow a$) naturally define units for mass, length, time,
and temperature.  (Planck remarked that such units would be meaningful
even in non-human and extraterrestrial cultures.)

A hypothetical Planckian SI unit system would have
\begin{equation} \label{eq:defs}
\begin{aligned}
\mathrm{kg} &= 4.594\ldots \times 10^7
               \left(\frac{\hbar c}{G}\right)^{\!1/2} \\
\mathrm{m}  &= 6.187\ldots \times 10^{34}
               \left(\frac{\hbar G}{c^3}\right)^{\!1/2} \\
\mathrm{s}  &= 1.854\ldots \times 10^{43}
               \left(\frac{\hbar G}{c^5}\right)^{\!1/2} \\
\mathrm{K}  &= 7.058\ldots \times 10^{-33}
               \left(\frac{\hbar c^5}{Gk_B^2}\right)^{\!1/2}
\end{aligned}
\end{equation}
with the numerical coefficients chosen so as to match the present
standards.  The square-rooted expressions are conventionally known as
the Planck mass, Planck length, and so on.  (They are actually
$\sqrt{2\pi}$ times Planck's definitions.)  Extension to electrical
units is straightforward, but we will not discuss it here.  The Planck
length, time, and temperature themselves are extreme scales, not
probed by any known physical process, and presumably the regime of
quantum gravity.  But extreme scales do not harm if they are only part
of some expression, as they are here.  The Planck mass is small but
microscopically measurable: $\sim20\mu\mathrm{g}$.  This is of the
same order as the uncertainty in the prototype kilogram itself.\cite{BIPM}  Hence
the kilogram could be redefined as an integer multiple of the Planck
mass, without significantly perturbing its value.

The difficulty with Eq.~\eqref{eq:defs} is that all the expressions
involve $G$.  Given the current uncertainties in $G$,\cite{CODATA16}
Eq.~\eqref{eq:defs} would allow units to be standardized only to four
significant digits, which is obviously not enough.  Improvement by
several orders of magnitude would be needed, to make the Planckian
kilogram practicable.  In this paper, we study a scenario which may
make such an improvement possible.  A Planckian second is not a prospect
with anything like present technology, because demands on the time
standard are more stringent still.\\

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{img/plotGmeas}
	\caption{Six of the most recent measurements of the gravitational constant $G$ with their uncertainties are plotted. The large disagreement is recognised immediately. The green lines represents the current recommended CODATA value and its uncertainty.}
	\label{fig:Gmeasurements}
\end{figure}

@@

In Fig.~\ref{fig:Gmeasurements} we have plotted some of the most recent measurements of the gravitational constant. Three of them have been carried out using a torsion balance [(BIPM-14)\cite{Quinn14}, (UCI-14)\cite{Newman14}, (HUST-09)\cite{Luo09}], (UZur-06)\cite{Schlamminger06} used a beam balance and (JILA-10)\cite{Parks10} measured the separation of two simple pendulums in a gravitational field. These measurements all tried to quantify the very weak gravitational force with a mechanical apparatus. The sixth measurement, (LENS-14)\cite{Rosi14}, was performed under another approach, using cold atom interferometry. Its uncertainty is a lot larger than the others, but it could become more competitive in the future. However, a brief glance on the plotted measurements shows the disagreement of current $G$-values and suggests a demand for new, potentially different techniques.\\

\subsection{Description of new measurement}
Newton's law of gravitation does not only explain why an apple falls down from a tree, it also explains orbits. If one has two bodies of masses $m_1$ and $m_2$, on which the only force acting is their respective gravitational attraction, they will orbit around their common barycenter in an ellipse described by Kepler's laws. The third law, given in Eq.~(\ref{eq:kepler3}), states that the square of the orbital period $P$ is directly proportional to the cube of the semimajor axis $a$. In this proportionality factor, the gravitational constant $G$ appears.
\begin{equation}\label{eq:kepler3}
P^2 = \frac{4 \, \pi^2}{G(m_1+m_2)}a^3
\end{equation}
If one is able to measure the position of the two masses very accurately and thus knows the orbit, it is possible to deduce the gravitational constant. 
So what we need is a freely falling two-body system. The only way to achieve this over some time, is to go to space. Thus, one has to design a spacecraft involving a cavity that contains the two masses and a position measurement apparatus. The cavity should shield the two-body system from environmental influences. The spacecraft is then put in space, preferably to a Lagrange point of the earth-sun system, in order to prevent large tidal forces.\\

@@

To get a feeling what the dimensions of such an experiment could be, choose for example lead balls with a diameter of about 10 cm, their mass then is about 6 kg. If one then chooses a semimajor axis of 15 cm, the orbital period would be roughly 3.5 hours and the gravitational force acting between the two bodies is of order $10^{-8}$ N. In Fig.~\ref{fig:sketch} we show a sketch on how this would look like. Treating this system with Newtonian gravity is justified, because corrections from general relativity would be of order $(v^2/c^2)$ and $(GM/rc^2)$, which for our two-body orbit are both very small numbers.\\

For getting an improvement in the precision of the gravitational constant, the positions of the two balls have to be known very accurately. Best suitable would probably be a laser interferometry system, which can yield precisions down to some nanometers.\cite{Loughridge13}
The main difference to other techniques is that position measurements can be very accurate compared to directly measuring the gravitational force.
This new kind of measurement of the gravitational constant appears simple, but it brings various difficulties with it. It is practically impossible to have a clean system, where only the respective gravitational force is acting between the two bodies, so one has to find out what effects could possibly disturb the system and then either make them as small as possible or correct for them. \\

@@

\textbf{Radiation pressure:} The envelope of the cavity needs to protect the two body-system from the constant bombardment of radiation and charged particles from the sun. Since the activity of the sun varies much, the spacecraft must be able to correct its position around the freely falling masses inside the cavity.\\
Furthermore, there is thermal radiation inside the cavity, that could disturb the orbit.
Using the Stefan-Boltzmann law, we can relate radiation pressure to the temperature inside the cavity. If we then specify to balls of radius $r$, we can express the force on the balls as 
\begin{equation}
F = \frac{8 \pi^6 r^2 {k_\mathrm{B}}^4 T^4}{45 c^3 h^3}.
\end{equation}
As we can see, this force is highly temperature dependent. In order to minimize this influence, one should deploy a temperature as small as possible in the cavity. In fact, if a cooling system based on liquid helium is installed, temperature can be brought down to about 4 K. Such a low temperature results in a force on the balls in the order of $10^{-16}$ N. Compared to the gravitational force, this is orders of magnitude less and thus we don't have to worry about effects of radiation pressure.\\

\textbf{Influence of the position measurement:} 
The photons of the laser that measures the position of the two balls exert an additional force. If a laser with a power of 1 mW is used and not more than five measurements every second are performed, the force on the balls is in the same order as the influence of thermal radiation and thus negligible too.\\

\textbf{Coulomb force:}
Since the electromagnetic force is orders of magnitude stronger than the gravitational force, the balls must be absolutely neutral. A possible source of charges are cosmic rays or the solar wind, so they have to be protected as good as possible. If we want that the Coulomb force between the balls is in the same order as the effect of thermal radiation and thus negligible, the maximal allowed charge on the balls is about $6 \cdot 10^{-14}\,\mathrm{C}$, which corresponds to roughly $4 \cdot 10^5$ elementary charges.\\

@@

\subsection{Orbital equation of motion}
From Newtonian dynamics the equation of motion for a clean two-body system is given by 
\begin{equation}
\label{eq:eomeasy}
\ddot{\vec{r}} = - \frac{G M\, \vec{r}}{r^3}.
\end{equation}
Here $M$ denotes the sum of the masses $m_1$ and $m_2$, $G$ is the gravitational constant and $r$ is the relative distance between the two bodies. For simplicity we will abbreviate the product of $GM$ with $k$.
Since our two-body system is in a rotating frame, we have to add the term developed before for the Coriolis acceleration as well as a correction term proportional to the distance accounting for centrifugal and tidal forces. This last term can be written with a
symmetric $3 \times 3$ matrix $C$ as follows  
\begin{equation}
\label{eq:tidal}
\ddot{\vec{r}}_{\mathrm{C}} = C \cdot \vec{r}.
\end{equation}
or in terms of the individual components 
\begin{equation}
\label{eq:tidacom}
\left(\ddot{r}_\mathrm{C}\right)_i = \sum_{k=1}^{3} C_{ik} \: r_k.
\end{equation}
It can be thought of as a first order expansion and therefore also include other possible minor perturbations. The reference frame is chosen such that the barycenter of the two-body system is force-free.
So the full equation of motion for the two balls looks as follows:
\begin{equation}
\label{eq:eom}
\ddot{\vec{r}} = - \frac{k\,\vec{r}}{r^3} - 2 \; \vec{\omega} \times \dot{\vec{r}} + C \cdot \vec{r}
\end{equation}
These are actually three coupled equations of motion, one for each coordinate axis. Since all are second order differential equations, this leaves us with six initial conditions. In other words to determine the orbit, we need to give six starting values.\\
We choose our coordinate system such that the orbital plane of the two-body system is in the $x-y$ plane. $\vec{\omega}$ is a constant vector that points in the direction of the rotation axis of the earth-sun reference frame. We will fit for $\vec{\omega}$, so we don't have to know it exactly before, but it can be estimated well and thus provides good starting points for the fit. The same is valid for the six components of the matrix $C$.\\

@@

Our problem has 16 unknown parameters, but the only one we are really interested in is $k$, which contains the gravitational constant. For the fit, we suggest to use a Markov Chain Monte-Carlo (MCMC) method because it does good work even if the parameter space is large. 
A Markov chain Monte-Carlo method is a Bayesian approach to data analysis. It uses the information provided by observed data about a set of parameters, to update a prior state of beliefs about the set of parameters to become a posterior state of beliefs about a set of parameters. We will not describe in detail how the MCMC is working, rather refer to an online tutorial from ESO PyCoffee\cite{esopycoffee} or an introduction paper from van Ravenzwaaij et al.\cite{vanRavenzwaaij16}

First we have to specify the prior probability distribution, so to say what values for the 16 parameters in the equation of motion actually make sense. Then, for all 16 parameters, a uniform distribution within these ranges and zero probability everywhere else is assumed as the prior probability distribution.
We briefly explain in what range the values of the parameters are estimated and display it in Table~\ref{tab:parameters}.
For the parameter $k$, which is the product of the gravitational constant $G$ and the total mass $M$, if we look at the current measured values for the gravitational constant in Fig.~\ref{fig:Gmeasurements}, it is reasonable to constrain its numerical value in SI units to values between $6.671 \cdot 10^{-11}$ and $6.676 \cdot 10^{-11}$. The masses of the two balls will not contribute to a larger uncertainty in $k$, we choose them to be 6 kg.

The vector $\vec{\omega}$ accounts for the rotation of the reference frame. If we say that the orbit of the balls lies in the same plane as the orbital plane of the earth, $\vec{\omega}$ will point in $z$ - direction with $x$ and $y$ components small. Thus $\omega_z$ is associated with the period $P_\odot$ of one year through $2\pi/P_\odot$. Because the rotation axis are conceivably not exactly aligned, we add some uncertainty. As mentioned before, the $x$ and $y$ components are small, so to say roughly in the range of the uncertainty in $w_z$.
The main contribution to the tidal term goes with $\omega^2$ and lies in the orbital plane. We are free to choose the $x$ and $y$ axis, so we choose it such that it is in $x$-direction. The other five components then should be smaller, but we give some extra range to account for possible other perturbations.
The initial values for the positions can be guessed very accurately, since they are directly measured precisely. The initial velocity can be calculated easily and matched to the initial positions, therefore also yielding narrow parameter ranges.

@@

In the next step, we set the parameters to a fixed ``true" value within the range estimated. We have to keep in mind that $k$, $x_0$ and $v_{y_0}$ are linked together and there must not be contradictions in the true values. The chosen values are displayed in the right column in Table~\ref{tab:parameters}.
To simulate data we would obtain from the position measurements, we integrate the equations of motion. The time for one orbit without perturbations can be estimated using Kepler's third law (\ref{eq:kepler3}). For the value of $k$ assumed above and a semi-major axis of $0.15 \, \mathrm{m}$, this results in an orbital period of roughly $12900 \, \mathrm{seconds}$ or $3.5 \, \mathrm{hours}$. Here we integrate two orbits in $100000$ steps, that would correspond to about $4$ measurements per second, which should be feasible. To the generated data, we add normal distributed noise with a standard deviation of $10 \, \mathrm{nm}$. This uncertainty stays for a systematic error in the laser interferometry measurement and also random statistic noise. Its value is just an estimation, it would depend highly on the measurement system that has been used.

@@

A useful way to look at the samples of the posterior returned by the MCMC is doing a corner plot using a Python module named \textit{corner}\cite{corner}. The corner plot shows all the one and two dimensional projections of the posterior probability distributions and demonstrates quickly the covariances between parameters. Along the diagonal, the marginalized distribution for each parameter independently is shown.
In Fig.~\ref{fig:plotCorner4} an excerpt of the full corner plot is shown.

To arrive at a best value with uncertainty for the gravitational constant, we take the mean and standard deviation of the marginalized distribution of the parameter $k$ and divide by the total mass of the balls. In this example the relative uncertainty of $k$ is roughly $10^{-7}$, so if the masses are measured to about the same precision, we can write for the gravitational constant $G$:
\begin{equation}\label{Gfit}
G = \left(6.674005 \pm 0.000001 \right) \cdot 10^{-11}  \frac{\mathrm{m}^3}{\mathrm{kg \, s}^2}
\end{equation}
This corresponds to a relative uncertainty of $1.5 \cdot 10^{-7}$ or 15 parts in a hundred million.

\section{Conclusion}
We have proposed a new way to measure the gravitational constant $G$ and estimated on what order of magnitude different disturbing effects occur and if one has to correct for them. We then simulated measurement data and showed how to extract the gravitational constant. Based on our estimations, it should be possible to measure the gravitational constant to about $10^{-7}$ relative uncertainty. This is a about a factor of 300 better than today's recommended CODATA value.

If we are a bit optimistic, a further increase in precision is not out of reach, so that about the same relative uncertainty as for Planck's constant $h$ ($1.2 \cdot 10^{-8}$)\cite{CODATA16} could be achieved. This then would allow to redefine the kilogram as an integer multiple of the Planck mass.\\

@@

@@

\bibitem{BIPM} Bureau International des Poids et Mesures, Calibration and measurement services: Mass, \url{<https://www.bipm.org/utils/common/pdf/cms_m.pdf>}, (2015).
